Computing Entropies With Nested Sampling

### 2017-03-13 paper/ms.tex
Introduction writing

The entropy quantifies the degree to which the issue
``what is the value of $x$, precisely?'' remains unresolved. For example,
if there is only one possible value which has probability 1, $H_x$ is
zero\footnote{Conventionally, $0 \times \log 0$ is defined to be equal
to $\lim_{x \to 0^+} (x\log x)$, which is zero.}.
If there are $N$ possibilities with equal probabilities $1/N$ each,
then the entropy is $H_x = \log(N)$.

@@

Throughout this paper, I use the compact
`overloaded' notation for probability distributions, writing $p(x)$ instead
of the more cumbersome $P(X=x)$.
Probability distributions are implicitly conditional on some prior
information, which is omitted from the notation unless necessary.
All logarithms are written $\log$ and any base can be used, unless otherwise
specified (for example by writing $\ln$ or $\log_{10}$). Any numerical
values given for the value of specific entropies are in {\em nats}, i.e.,
the natural logarithm was used.

@@

@@

\subsection{The relevance of data}
The degree to which the question ``what is the value of $\x$?'' is
answered by the question ``what is the value of $\d$?'' is given by the
conditional entropy
\begin{align}
H_{\x | \d} &= - \sum_{\d} p(\d) \sum_{\x} p(\x | \d) \log p(\x | \d)
\end{align}
which is the expected value of the entropy of the posterior, averaged over
all possible datasets which might be observed.
This is difficult to compute because the expression
for the posterior distribution
\begin{align}
p(\x | \d) &= \frac{p(\x)p(\d | \x)}{p(\d)}
\end{align}
contains the marginal likelihood $p(\d)$.

>>>>

The entropy quantifies the degree to which the issue
``what is the value of $x$, precisely?'' remains unresolved. For example,
if there is only one possible value which has probability 1, $H_x$ is
zero\footnote{Conventionally, $0 \times \log 0$ is defined to be equal
to $\lim_{x \to 0^+} (x\log x)$, which is zero.}.
If there are $N$ possibilities with equal probabilities $1/N$ each,
then the entropy is $H_x = \log(N)$.
If the space of possible $x$-values is continuous, the differential
entropy
\begin{align}
H_{x} &= -\int p(x) \log p(x) \, dx
\end{align}
quantifies uncertainty by generalising the log-volume
of the plausible region, where volume is defined with respect to $dx$.

@@

Throughout this paper, I use the compact
`overloaded' notation for probability distributions favoured by
Bayesians, writing $p(x)$ (for either a probability mass function
or a probability density function) instead
of the more correct but cumbersome $P(X=x)$ or $f_X(x)$.
Probability distributions are implicitly conditional on some prior
information, which is omitted from the notation unless necessary.
All logarithms are written $\log$ and any base can be used, unless otherwise
specified (for example by writing $\ln$ or $\log_{10}$). Any numerical
values given for the value of specific entropies are in {\em nats}, i.e.,
the natural logarithm was used.

@@

In practical applications of Bayesian inference, the following operations
are usually feasible, and are assumed to be possible at low computational
cost:
\begin{enumerate}
\item Points can be generated from the prior $p(\x)$;
\item Simulated datasets can be generated from $p(\d | \x)$ for a given value of $\x$;
\item The likelihood, $p(\d | \x)$ can be evaluated cheaply for any $\d$ and $\x$.
\end{enumerate}

\subsection{The relevance of data}

@@

The degree to which the question ``what is the value of $\x$?'' is
answered by the question ``what is the value of $\d$?'' is given by the
conditional entropy
\begin{align}
H_{\x | \d} &= - \sum_{\d} p(\d) \sum_{\x} p(\x | \d) \log p(\x | \d)
\end{align}
which is the expected value of the entropy of the posterior, averaged over
all possible datasets which might be observed.

This is difficult to compute because the expression
for the posterior distribution
\begin{align}
p(\x | \d) &= \frac{p(\x)p(\d | \x)}{p(\d)}
\end{align}
contains the marginal likelihood $p(\d)$.
